An Invariant Framework for Conducting Reproducible Computational Science

### 2015-01-14 intro.tex
Docker changes

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated and executable records of computations, especially computations whose results are destined for publication and/or re-use, and the VMIs can be shared easily \cite{}. 
With this approach, oftentimes, the resulting image is too big in size for share-ability purposes. An alternative light-weight form of virtualization allows encapsulation of the application software, along with all its necessary dependencies into a self-contained package, but does not make it executable.  It thus separates, and does not include the underlying operating system on which the package is executed. The encapsulation of the self-contained package is achieved by interposing application system calls, and only copying the necessary dependencies (data, libraries, code) into the package, which also makes it light-weight than a VMI. While both forms of virtualization provide mechanisms for sandboxing the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result.

Since reproducibility is significantly about comprehensive documentation, virtualization approaches in their current form only make it easy to capture the computations. Preserving the computations, so that they are easy to understand, install, or alter implicit dependencies that are part of computation, especially as dependencies and software components evolve or become deprecated, is not effectively addressed. There are two approaches to address the preservation challenge. By either introducing tools that help document dependencies and provide software attribution within VMIs or packages, or alternatively by using software delivery mechanisms, such as, centralized package management, linux containers and Docker framework. We examined the first approach previously in \cite{}. In this paper we examine the the second approach, that is, examine how the documentation and preservation standards that are part of software delivery mechanisms can be combined with traditional virtualization approaches for reproducibility. In particular we consider the light-weight virtualization approaches, because, we believe that the combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers. 
% they may help the This  We that must also include effective  to deliver the software to a wide audience. We examine the fire
%have recently emerged  several of the reproducible challenges that arise due to use of virtualization approaches.

To conduct a thorough examination, we consider a real-world complex high energy physics application, independently developed by a group, but must be reproduced  so that the entire high energy physics community can benefit from the analysis obtained from the application. The collaborative nature of in which high energy physics science is conducted is a 
We describe challenges in reproducing this application, and consider the extent to which reproducibility can be satisfied, independently with light-weight virtualization approaches and software delivery mechanisms. We then propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms. The framework uses light-weight virtualization for capturing a computational environment, but preserves the environment using a software delivery approach, so it can be re-executed later. By using standardized delivery mechanisms and expanding the application for future uses, the framework makes it invariant to temporal software changes that may occur in near future. We measure the performance overhead of combining light-weight virtualization with software delivery approaches. The combination imposes additional overhead due to double virtualization, but saves on context switches, since every system call is not interposed.

>>>>

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated, and executable records of computations, especially computations whose results are destined for publication and/or re-use, and the VMIs can be shared easily \cite{}. 
With this approach, oftentimes, the resulting image is too big in size for share-ability purposes. An alternative light-weight form of virtualization allows encapsulation of the application software, along with all its necessary dependencies into a self-contained package, but does not make it executable.  It, thus, separates and does not include the underlying operating system on which the package is executed. The encapsulation of the self-contained package is achieved by interposing application system calls, and only copying the necessary dependencies (data, libraries, code, etc) into the package, which also makes it lighter weight than a VMI. While both forms of virtualization provide mechanisms for sandboxing the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result.

Since reproducibility is significantly about comprehensive documentation, virtualization approaches in their current form only make it easy to capture the computations. Preserving the computations, so that they are easy to understand, install, or alter implicit dependencies that are part of computation, especially as dependencies and software components evolve or become deprecated, is not effectively addressed. There are two approaches to address the preservation challenge. By either introducing tools that help document dependencies and provide software attribution within VMIs or packages, or alternatively by using software delivery mechanisms, such as, centralized package management, linux containers, and the, more recent, Docker framework. We examined the first approach previously in \cite{}. In this paper we examine the the second approach, that is, examine how the documentation and preservation standards that are part of software delivery mechanisms can be combined with traditional virtualization approaches for reproducibility. In particular, we consider the light-weight virtualization approaches, because, we believe that the combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers. 
% they may help the This  We that must also include effective  to deliver the software to a wide audience. We examine the fire
%have recently emerged  several of the reproducible challenges that arise due to use of virtualization approaches.

To conduct a thorough examination, we consider real-world complex high energy physics (HEP) applications, independently developed by two HEP groups, that must be reproduced so that the entire high energy physics community can benefit from the analysis obtained from the application. We describe challenges in reproducing the applications, and consider the extent to which reproducibility requirements can be satisfied with light-weight virtualization approaches and software delivery mechanisms. We propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms for efficiently capturing, invariantly preserving and practically deploying applications. In effect, the framework uses light-weight virtualization techniques for efficiently capturing a computational environment, but preserves the environment such that it is invariant to temporal software changes that may occur in near future and can be re-executed later. We measure the performance overhead of  light-weight virtualization and software delivery approaches. 
% For later. 
%The combination imposes additional overhead due to double virtualization, but saves on context switches, since every system call is not interposed.

### 2015-01-14 measure.tex
Add preservation_framework figure
Docker changes

% PTU
\vspace{5pt}
\noindent{\bf CDE} CDE creates a package of an application by recording program binaries, libraries, scripts, data files, and environment variables, etc. It also uses the UNIX ptrace system call interposition to identify the code and data used by the running application. The original CDE as available through [13,12] was limited to major Linux kernel versions (e.g. 2.6.x), but we have removed that restriction by adapting it for the newly released Linux kernel 3.0 as well as for Mac OS X \cite{}. Thus CDE, unlike Parrot, provides OS invariance and does not require installation and configuration.

@@

% PTU package
\vspace{5pt}
\noindent{\bf CDE:} In CDE, the complete package is preserved, consisting of all code and data files, however deep the data path is present. 
However, a CDE binary only captures a single execution path, which is the execution path taken during run-time. If different execution paths need different types of dependencies, some dependencies may be left out. To preserve these dependencies, we have added addition functionality to CDE. In particular, information about other static binaries and required static shared libraries, such as version number, released version of shared libraries, and associated kernel distribution, is preserved using UNIX commands file, ldd, strings, and objdump. We also have added a provenance wrapper over CDE. This provenance wrapper documents all the dependencies, and provides information about process execution times and memory consumption. It creates a graph of the reference execution providing details about how data was created and transformed as part of the execution. Thus instead of a PLU, a description of the PLU exists in the package.

Neither the Parrot packages or extended CDE package provide a literate programming environment or a continuous integration environment for changes in software. Thus they are partially preserved but not completely.

% Docker images

\subsection{Deployment  Repositories}
image can be stored and distributed in a repository like Docker Hub. 
The Hub also provides a convenient dis- tribution service, freely storing the pre-built images, along with their metadata, for download and reuse by others. 
   The Docker Hub is a free service and an open source software product so that users can run their own private versions of the Hub on their own servers, for instance, if security of the data or the longevity of the public platform is a concern.

>>>>

% PTU
\vspace{5pt}
\noindent{\bf PTU} PTU creates a package of an application by recording program binaries, libraries, scripts, data files, and environment variables, etc. It also uses the UNIX \emph{ptrace} system call interposition to identify the code and data used by the running application. PTU was forked from CDE \cite{}, and was thus limited to major Linux kernel versions (e.g. 2.6.x); consequently we have removed that restriction by adapting it for the newly released Linux kernel 3.0 as well as for Mac OS X \cite{}. Thus PTU provides OS invariance and does not require installation and configuration.

@@

% PTU package
\vspace{5pt}
\noindent{\bf PTU:} Owing to CDE, in PTU, the complete package is preserved, consisting of all code and data files, however deep the data paths are. 
In CDE a binary only captures a single execution path, which is the execution path taken during run-time. If different execution paths need different types of dependencies, some dependencies may be left out. To preserve these dependencies, we have added addition functionality in PTU. In particular, information about other static binaries and required static shared libraries, such as version number, released version of shared libraries, and associated kernel distribution, is preserved using UNIX commands file, ldd, strings, and objdump \cite{}. We also have added a provenance wrapper over CDE \cite{}. This provenance wrapper documents all the dependencies, and provides information about process execution times and memory consumption. It creates a graph of the reference execution providing details about how data was created and transformed as part of the execution. The proevenance description is analgalous to the logically preserved unit.

Neither the Parrot packages or extended CDE package provide a literate programming environment or a continuous integration environment for changes in software. Thus they are preserved but make break over time as technology changes.

 % Docker images
\vspace{5pt}
\noindent{\bf Linux Containers and Docker Images}

Linux Containers provide multiple isolated execution instances on top of the same kernel through OS-level virtualization. Thus they can be used to persist the captured packages into images. Using such containers, Docker now allows to preserve the image in a more user-friendly way. In particular, apart from performing Linux container (LXC) based operating system (OS) level virtualization, docker tools provides portable deployment of containers across platforms, documentation of packages in a scriptable format, and versioning of container images. The image can be preserved along with dockerfiles, which are plain text files with a literate programming environment: the computational part is similar to shell scripts and can help provision similar to other provisioning tools (e.g. Chef, Puppet) or Continuous Integration (CI) platforms (e.g. Travis CI), and the text part is for human consumption and more suited for use with a version management system such as subversion or git, which can track any changes made to the Dockerfile. Thus dockerfiles can be used to preserve the namelist of Parrot packages and provenance description in PTU. 
Docker is integrated with a continuous build environment which will check and validate the version of the software being used, and use a more recent version to build application software.

\subsection{Deployment  Repositories}
Deployment repositories, such as Docker Hub, are distribution service for storing the pre-built images, along with their metadata, for download and reuse by others. While the architecture of such hubs is beyond the scope of the current paper, the existence of such a repository for reproducible research is vital to ensure security and longevity of preserved packages.

### 2015-01-14 paper.tex
Modify Acknowledgement Section; Remove Postscript Section
Docker changes

\author{
	Haiyan Meng\inst{2}
\and Rupa Kommineni\inst{1}
\and Quan Pham\inst{1} \\
\and Robert Gardner\inst{1}
\and Tanu Malik\inst{1}
\and
	Douglas Thain\inst{2}
}
\institute{
	Computation Institute,
	University of Chicago,
	Chicago, Illinois, USA \\
	\email{rupa, quan, rob, tanum@uchicago.edu}
\and
	Department of Computer Science and Engineering,
	University of Notre Dame,
	Notre Dame, Indiana, USA \\
	\email{hmeng, dthain@nd.edu}
}

@@

\authorrunning{Meng et al.}

@@

\begin{abstract}
\it Computational reproducibility depends on being able to isolate necessary and sufficient computational artifacts and preserve them for later re-execution.
Both isolation and preservation of artifacts can be challenging due to the complexity
of existing software and systems and the resulting implicit dependencies, resource distribution, and shifting compatibility of systems as time progresses---all conspiring
to break the reproducibility of an application. Sandboxing is a technique
that has been used extensively in OS environments for isolation of computational artifacts.
Several tools were proposed recently that employ sandboxing as a mechanism to ensure reproducibility.
However, none of these tools preserve the sandboxed application for re-distribution
to a larger scientific community---aspects that are equally crucial for ensuring reproducibility as sandboxing itself.
In this paper, we describe a combined sandboxing and preservation framework, which is efficient and practical and
useful for large-scale reproducibility. We present a case study of a complex high energy
physics application and show how the framework can be useful for sandboxing and preserving.
We report on the completeness, performance,
and efficiency of the framework, and suggest possible standardization approaches.
\end{abstract}

>>>>

\author{
	Haiyan Meng\inst{2}
\and Rupa Kommineni\inst{1}
\and Quan Pham\inst{1} \\
\and Robert Gardner\inst{1}
\and Tanu Malik\inst{1}
\and
	Douglas Thain\inst{2}
}
\institute{
	Computation Institute,
	University of Chicago,
	Chicago, Illinois, USA \\
	\email{rupa, quanpt, rwg, tanum@uchicago.edu}
\and
	Department of Computer Science and Engineering,
	University of Notre Dame,
	Notre Dame, Indiana, USA \\
	\email{hmeng, dthain@nd.edu}
}

@@

\authorrunning{Meng, and Thain}

@@

\begin{abstract}
\it Computational reproducibility depends on being able to isolate necessary and sufficient computational artifacts and preserve them for later re-execution.
Both isolation and preservation of artifacts can be challenging due to the complexity
of existing software and systems and the resulting implicit dependencies, resource distribution, and shifting compatibility of systems as time progresses---all conspiring
to break the reproducibility of an application. Sandboxing is a technique
that has been used extensively in OS environments for isolation of computational artifacts.
Several tools were proposed recently that employ sandboxing as a mechanism to ensure reproducibility.
However, none of these tools preserve the sandboxed application for re-distribution
to a larger scientific community---aspects that are equally crucial for ensuring reproducibility as sandboxing itself.
In this paper, we describe a combined sandboxing and preservation framework, which is efficient, invariant and
practical for large-scale reproducibility. We present case studies of complex high energy
physics applications and show how the framework can be useful for sandboxing, preserving and distributing applicatios.
We report on the completeness, performance,
and efficiency of the framework, and suggest possible standardization approaches.
\end{abstract}
